A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2017-07-11 intro.tex
incorporating edits in intro + flagship question

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economic systems, land use, the climate,
and air pollution. IAMs provide global projections of
systemic change by dividing the world into a number of representative regions
(typically 10 to 30), the definition of which is distinct for each model
\cite{krey_global_2014}. Results from IAMs are integral in a number of
international studies, which notably include projections of climate and economic
futures. Recently, flagship IAMs have developed scenarios based on the Shared
% * <ssmith@pnnl.gov> 2017-07-01T03:11:26.621Z:
% 
% > flagship
% Suggest different word. "comprehensive" is one possibility.
% 
% ^ <ssmith@pnnl.gov> 2017-07-01T03:11:57.571Z.
Socioeconomic Pathways (SSPs) \cite{van_vuuren_energy_2017, fricko_marker_2017,
  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
quantify a variety of potential global futures. The SSPs are  designed
to be used in  research that include earth system
model (ESM) simulations, climate impact, adaptation and climate mitigation
studies.

While IAMs are implemented in myriad ways, including simulation and
optimization, the core inputs, and outputs  are similar across different
models. Modeling teams incorporate  data on energy systems, land
use, economics, demographics and emissions sources and concentrations, among other
data, in order to provide consistent future trajectories of modeled
variables. The models then provide estimates of future trajectories of these
variables under various socioeconomic and technological assumptions as well as
proposed policy constraints, e.g., targets for future Greenhouse Gas (GHG)
emissions.

The emissions trajectories calculated by IAMs are critical inputs for ongoing,
worldwide scientific community efforts in the Coupled Model Intercomparison
Project Phase 6 (CMIP6) \cite{eyring_overview_2016}, which is utilizing a number
of marker SSP scenarios developed by the IAM community
\cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
modeling the individual technologies and sectors that contribute towards the
emissions of different air pollutants and GHGs as well as various mitigation
technologies (e.g., NOx and SOx scrubbers on power plants). However, the
historical emissions starting points of  models can differ by
large amounts depending on the region, sector, and emissions species.

@@

The remainder of this paper describes the methodology and implementation of the
\code{aneris} code base \cite{matthew_gidden_2017_802832}, written in the Python
programming language (detailed documentation available
\href{http://mattgidden.com/aneris/}{online}). Section \ref{sec:meths} provides
a detailed description of the underlying mathematical components of
\code{aneris} as well as the code workflow. The results of applying the
automated harmonization mechanism on two example IAM scenarios, one with
emissions growth and another with emissions mitigation, is presented in Section
\ref{sec:results}. Finally, the general effectiveness and potential future
improvements on the automated methodology is discussed in Section
\ref{sec:future}.

% todo: add?
% what aspects of harmonization are important (e.g., cumulative carbon emissions).

>>>>

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economic systems, land use, the climate,
and air pollution. IAMs provide global projections of
systemic change by dividing the world into a number of representative regions
(typically 10 to 30), the definition of which is distinct for each model
\cite{krey_global_2014}. Results from IAMs are integral in a number of
international studies, which notably include projections of climate and economic
futures. Recently, the IAM community has developed scenarios based on the Shared
% * <ssmith@pnnl.gov> 2017-07-01T03:11:26.621Z:
% 
% > flagship
% Suggest different word. "comprehensive" is one possibility.
% 
% ^ <ssmith@pnnl.gov> 2017-07-01T03:11:57.571Z.
Socioeconomic Pathways (SSPs) \cite{van_vuuren_energy_2017, fricko_marker_2017,
  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
quantify a variety of potential global futures. The SSPs are  designed
to be used in  research that include earth system
model (ESM) simulations, climate impact, adaptation and climate mitigation
studies.

While IAMs are implemented in myriad ways, including simulation and
optimization, the core inputs and outputs  are similar across different
models. Modeling teams incorporate  data on energy systems, land
use, economics, demographics and emissions sources and concentrations, among other
data, in order to provide consistent existing trajectories of modeled
variables. The models then provide estimates of future trajectories of these
variables under various socioeconomic and technological assumptions as well as
proposed policy constraints, e.g., targets for future Greenhouse Gas (GHG)
emissions.

The emissions trajectories calculated by IAMs are critical inputs for ongoing,
worldwide scientific community efforts in the Coupled Model Intercomparison
Project Phase 6 (CMIP6) \cite{eyring_overview_2016}, which is utilizing a number
of marker SSP scenarios developed by the IAM community
\cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
modeling the individual technologies and sectors that contribute towards the
emissions of different air pollutants and GHGs as well as various mitigation
technologies. However, the historical emissions starting points of  models can differ by
large amounts depending on the region, sector, and emissions species.

@@

The remainder of this paper describes the methodology and implementation of the
harmonization code base \code{aneris} \cite{matthew_gidden_2017_802832}, written in the Python
% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:12:05.654Z:
% 
% > the
% > \code{aneris} code base \cite{matthew_gidden_2017_802832
% perhaps something like: "the harmonization code base \code{aneris}", to be explicit.
% 
% ^.
programming language (detailed documentation available
\href{http://mattgidden.com/aneris/}{online}). Section \ref{sec:meths} provides
a detailed description of the underlying mathematical components of
\code{aneris} as well as the code workflow. The results of applying the
automated harmonization mechanism on two example IAM scenarios, one with
emissions growth and another with emissions mitigation, is presented in Section
\ref{sec:results}. Finally, the general effectiveness and potential future
improvements on the automated methodology is discussed in Section
\ref{sec:future}.

### 2017-07-11 meth.tex
incorporating edits in intro + flagship question

The value of $c_v$ is then tested against a threshold. To determine this
threshold, an analysis of the recent CEDS and LUC historical data has been
performed. Figure \ref{fig:cov} shows the distribution of LUC $c_v$s and non-LUC
$c_v$s as determined for historical data aggregated to the model regions of 5
different flagship IAMs. A threshold value of 20 has been chosen based on these
observations as it optimally divides the two distributions. Importantly, tails
of the LUC and non-LUC overlap, thus there are both false positives (~10\% of
non-LUC trajectories) and false negatives (~7\% of LUC trajectories). However,
% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:07:21.157Z:
% 
% >  (~10\% of
% > non-LUC trajectories) and false negatives (~7\% of LUC trajectories)
% visually it seems to be that the false negatives are a larger share thatn the false positives?
% 
% ^.
as any regional definition is model dependent and thus any regional aggregation
is possible an automated detection mechanism is necessary.

@@

Finally, consideration is taken with respect to the relative difference between
the historic and model values in the harmonization time period. In order to
investigate the possible values that these relative differences can take, the
flagship IAMs values used in the SSP and (ongoing) CMIP6 inter-comparison
exercises are used. A distribution of these differences for all models in the
study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
value of 50\% was chosen to be used as a default in \code{aneris}.
% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:25:47.452Z:
% 
% > 50\%
% What does this mean? I would expect something of the form low_value < dH < high_value?
% 
% ^.

>>>>

The value of $c_v$ is then tested against a threshold. To determine this
threshold, an analysis of the recent CEDS and LUC historical data has been
performed. Figure \ref{fig:cov} shows the distribution of LUC $c_v$s and non-LUC
$c_v$s as determined for historical data aggregated to the model regions of 5
different IAMs \TODO{list here or elsewhere?}. A threshold value of 20 has been chosen based on these
observations as it optimally divides the two distributions. Importantly, tails
of the LUC and non-LUC overlap, thus there are both false positives (~10\% of
non-LUC trajectories) and false negatives (~7\% of LUC trajectories). However,
% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:07:21.157Z:
% 
% >  (~10\% of
% > non-LUC trajectories) and false negatives (~7\% of LUC trajectories)
% visually it seems to be that the false negatives are a larger share thatn the false positives?
% 
% ^.
as any regional definition is model dependent and thus any regional aggregation
is possible an automated detection mechanism is necessary.

@@

Finally, consideration is taken with respect to the relative difference between
the historic and model values in the harmonization time period. In order to
investigate the possible values that these relative differences can take, the
IAM values used in the SSP and (ongoing) CMIP6 inter-comparison
exercises are used. A distribution of these differences for all models in the
study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
value of 50\% was chosen to be used as a default in \code{aneris}.
% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:25:47.452Z:
% 
% > 50\%
% What does this mean? I would expect something of the form low_value < dH < high_value?
% 
% ^.
